---
title: "Data Assimilation for WRF Numerical Weather Prediction Outputs"
author: "Mehmet AKSOY & Sercan AKIL"
format: html
editor: visual
---

## Introduction

### WRF Numerical Weather Prediction Model

### Data Assimilation

## Implementation in R

There are two files for two different domains as assimilated and non-assimilated predictions of WRF model. Thus, there are four netcdf files totally. We are going to define prediction variables, forecast period and coverage of domains for each file.

### Uploading Necessary Packages

We will need several packages for some implementations in R, for instance; opening of the netcdf files of WRF data, handling of WRF outputs, visualization and etc.

```{r}
#| warning: false
library(tidyverse)
library(raster)
library(rnaturalearth)
library(ncdf4)
library(R.utils)
library(sf)
library(gt)
```

### Assimilated WRF Data

Firstly, we opened the assimilated WRF output for domain 1. You can see the content of WRF output below, the data cover time and coverage domain information with meteorological predictions such as wind speed **(U10 & V10)**, precipitation **(RAINC & RAINNC)**, etc. We need to use name of the variables to extract specific data variables from the raw data.

```{r}
fname <- paste0("D:/Kitaplar/METU-PHD/Thesis/IsmailHocandanAldim_Aksoy_27092023/",
"aswout/wrfout_d01_2004-08-11_00_00_00")
nc_data <- nc_open(fname)
```

### Precipitation Prediction (Domain 1)

In the WRF data, **RAINC** is defined as *ACCUMULATED TOTAL CUMULUS PRECIPITATION* which is convective precipitation and **RAINNC** is *ACCUMULATED TOTAL GRID SCALE PRECIPITATION* which is non-convective precipitation. To obtain the total precipitation, we just need sum these two variables: **PRECIPITATION = RAINC + RAINNC**.

```         
    float RAINC[west_east,south_north,Time]   
        FieldType: 104
        MemoryOrder: XY 
        description: ACCUMULATED TOTAL CUMULUS PRECIPITATION
        units: mm
        stagger: 
        coordinates: XLONG XLAT
```

```         
  float RAINNC[west_east,south_north,Time]   
        FieldType: 104
        MemoryOrder: XY 
        description: ACCUMULATED TOTAL GRID SCALE PRECIPITATION
        units: mm
        stagger: 
        coordinates: XLONG XLAT
```

### Spatial Resolution (Domain 1)

Spatial resolution of domain 1 is 12 km as you can see below *(DX: 12000, DY:12000)*.

```         
78 global attributes:
    TITLE:  OUTPUT FROM WRF V3.1.1 MODEL
    START_DATE: 2004-08-11_00:00:00
    SIMULATION_START_DATE: 2004-08-11_00:00:00
    WEST-EAST_GRID_DIMENSION: 192
    SOUTH-NORTH_GRID_DIMENSION: 116
    BOTTOM-TOP_GRID_DIMENSION: 28
    DX: 12000
    DY: 12000
```

There are **41** time steps which means we can get convective, non-convective and total precipitation predictions along the domain **(191x115)**. Additionally, the data includes information for **27** layers from bottom to top of the atmosphere. However, for the precipitation data we only need to calculate the precipitation that will reach the ground.

```{r}
long<- ncvar_get(nc_data, "XLONG")
lat<- ncvar_get(nc_data, "XLAT", verbose = F)
rain_c <- ncvar_get(nc_data, "RAINC") 
rain_nc <- ncvar_get(nc_data, "RAINNC") 

dim(rain_c);dim(rain_nc)
```

### Forecast Period & Time Interval (Domain 1)

We can also obtained the forecast horizon by getting time steps from the data. After getting time steps, we see that forecast period is between 11 and 16 (00:00 UTC) August, 2004. In this case, the time interval for the forecast period up to +120 hours *(or five days)* is three hour.

```{r}
t <- ncvar_get(nc_data, "Times"); t
ymd_hms(t[41]) - ymd_hms(t[1])
```

### Study Area (Domain 1)

@fig-domain1 below shows the coverage of domain 1 where covers Türkiye and its surrounding.

```{r}
#| label: fig-domain1
#| fig-cap: 
#|  Türkiye with its neighboring countries and coverage of domain 1 for assimilated-WRF predictions.

raster_rain_c<- list()
for (i in 1:dim(rain_c)[3])   {
    raster_rain_c[[i]] <- raster(t(rain_c[, , i]), 
       xmn=min(long), xmx=max(long),
       ymn=min(lat), ymx=max(lat), 
       crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
                              }

rain_c_df <-as.data.frame(raster_rain_c[[length(t)]], xy = TRUE) 
world <- rnaturalearth::ne_countries(scale='medium',returnclass = 'sf')


ggplot(data = world)  + geom_sf(fill = "white") + 
  coord_sf(crs = st_crs(4326), xlim = c(10, 55), ylim = c(30,50)) +
  geom_raster(data = rain_c_df, aes(x = x, y = y, fill = layer), 
            alpha=0.7, show.legend = FALSE) +
  scale_fill_viridis_c() +  labs(x="",y="") + 
  ggtitle("Coverage of Domain 1") 
```

### Precipitation Prediction (Domain 2)

There are **121** time steps for second domain since we can inference by dimension of **RAINC** data.

```{r}
fname2 <- paste0("D:/Kitaplar/METU-PHD/Thesis/IsmailHocandanAldim_Aksoy_27092023/",
"aswout/wrfout_d02_2004-08-11_00_00_00")
nc_data2 <- nc_open(fname2)

long_2<- ncvar_get(nc_data2, "XLONG")
lat_2<- ncvar_get(nc_data2, "XLAT", verbose = F)
rain_c_2 <- ncvar_get(nc_data2, "RAINC") 
rain_nc_2 <- ncvar_get(nc_data2, "RAINNC") 

dim(rain_c_2);dim(rain_nc_2)
```

### Spatial Resolution (Domain 2)

Spatial resolution of domain 2 is 4 km which is a finer resolution than previous one *(DX: 4000, DY: 4000)*.

```         
78 global attributes:
    TITLE:  OUTPUT FROM WRF V3.1.1 MODEL
    START_DATE: 2004-08-11_00:00:00
    SIMULATION_START_DATE: 2004-08-11_00:00:00
    WEST-EAST_GRID_DIMENSION: 133
    SOUTH-NORTH_GRID_DIMENSION: 64
    BOTTOM-TOP_GRID_DIMENSION: 28
    DX: 4000
    DY: 4000
```

### Forecast Period & Time Interval (Domain 2)

Forecast period for domain 2 is same with previous one. However, the time interval is one hour and it has a finer temporal resolution.

```{r}
t2 <- ncvar_get(nc_data2, "Times"); t2
ymd_hms(t2[121]) - ymd_hms(t2[1])
```

### Study Area (Domain 2)

@fig-domain2 below shows the comparison of two domains and coverage of domain 2 where covers some part of northwest of Türkiye.

```{r}
#| warning: false
#| label: fig-domain2
#| fig-cap: Comparison of domain 1 and 2. 

raster_rain_c_2<- list()
for (i in 1:dim(rain_c_2)[3])   {
    raster_rain_c_2[[i]] <- raster(t(rain_c_2[, , i]), 
       xmn=min(long_2), xmx=max(long_2),
       ymn=min(lat_2), ymx=max(lat_2), 
       crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
                              }

rain_c_df_2 <-as.data.frame(raster_rain_c_2[[length(t2)]], xy = TRUE) 
world <- rnaturalearth::ne_countries(scale='medium',returnclass = 'sf')

ggplot(data = world)  + geom_sf(fill = "white") + 
  coord_sf(crs = st_crs(4326), xlim = c(19, 47.5), ylim = c(33.5,47)) +
  geom_raster(data = rain_c_df, 
            aes(x = x, y = y, fill = layer), alpha=0.4, show.legend = FALSE) + 
  geom_raster(data = rain_c_df_2, 
            aes(x = x, y = y, fill = layer), alpha=0.7, show.legend = FALSE) + 
  scale_fill_viridis_c() + labs(x="",y="") + 
  ggtitle("Coverage of Domain 2") 
```

```{r}
turkey <- subset(world, admin == "Turkey")

ggplot(data = turkey)  + geom_sf(fill = "white") + 
  coord_sf(crs = st_crs(4326), xlim = c(extent(rain_c_df_2)[1], extent(rain_c_df_2)[2]), ylim = c(extent(rain_c_df_2)[3],extent(rain_c_df_2)[4])) +
 
  geom_raster(data = rain_c_df_2, 
            aes(x = x, y = y, fill = layer), alpha=0.7, show.legend = FALSE) + 
  scale_fill_viridis_c() + labs(x="",y="") + 
  ggtitle("Coverage of Domain 2") 

extent(rain_c_df_2)
```

@tbl-gauges-domain2 shows the main gauges, only one station was selected for each province, across the domain 2.

```{r}
#| label: tbl-gauges-domain2
#| tbl-cap: Stations accross the domain 2.

df_gauges <-  read.delim("D:/Kitaplar/METU-PHD/COURSES/3-TERM/STAT570/STAT_570_FINAL_PROJECT_MAKSOY-SAKIL/gauges.txt", sep="|")
df_gauges<- df_gauges[,-c(3,4)]
colnames(df_gauges)<- c("Station","Province","Latitude","Longitude","Altitude") ; 

df_gauges$Province <- tolower(df_gauges$Province) |> str_to_title() 

df_gauges |> gt()


```
